Synthesizing Quotient Lenses

### 2017-10-06 papers/QuotientLensSynth/main.tex
a few comments

\newcommand{\FINISH}[3]{\ifdraft\textcolor{#1}{[#2: #3]}\fi}
\newcommand{\bcp}[1]{\FINISH{dkred}{B}{#1}}
\newcommand{\BCP}[1]{\FINISH{dkred}{B}{\bf #1}}
\newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}}
\newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
\newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
\newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}

@@

@@

\saz{It might be a good idea to show what this would look like in all its gory
detail, just to make the problem very concrete. Perhaps in a figure?}\bcp{+1}

In this paper, we introduce the formalism of Quotient Regular Expressions (or
QREs) to address this issue. QREs enable the programmer to express
a rich set of equivalence relations on regular expressions at the level of
syntax. QREs therefore enable the programmer to express more equivalence
relations with less effort. Also, since QREs are defined at the level of
syntax on regular expressions, QREs give the programmer more control over where
the quotienting occurs.

@@

@@

  \begin{enumerate}
    \item 
    Every regular expression can be lifted to a QRE. When a regular expression
    $R$ is used as a QRE, then $\mathcal{L}(R)$ is assumed to be quotiented by
    the equality equivalence relation on $\mathcal{L}(R)$.
    \item
    For the QRE $R \mapsto s$, the data is described by $\mathcal{L}(R)$, but
    all members of $R$ are considered equivalent to each other, with $s \in
    \mathcal{L}(R)$ as the choice of representative. This construct is useful
    when the programmer simply wishes to ignore unnecessary characters such as
    excess whitespace e.g. 
    $$\mathtt{FIRST\_NAME} \cdot (\mathtt{WHITESPACE}^* \mapsto `` \quad ")
    \cdot \mathtt{LAST\_NAME}$$
    \item
    $\squash{R}{R'}{f}$ enables the programmer to use one of two different
    representations of the data if they can provide a way of mapping the first
    representation to the second. For instance, the first representation may be
    $$\mathtt{FIRST\_NAME} \cdot `` \quad " \cdot \mathtt{LAST\_NAME}$$
    while the second representation may be
    $$\mathtt{FIRST\_NAME} \cdot ``," \cdot \mathtt{LAST\_NAME}$$
    The mapping that maps the first representation to the second simply converts
    the space between the first and last name to a comma.
    \item
    The $\perm{q_1, \ldots, q_n}{q}$ construct enables the programmer to
    consider all permutations of some data that are interspersed with a
    separator the same. For instance, in the \textsc{Bib}\TeX{} to EndNote
    example, the programmer may use this construct as follows:
    $$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
    \re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
 where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$ are
regular expressions describing the respective \textsc{Bib}\TeX\; fields of the
same name, and $\re{WHITESPACE}$ is a regular expression describing whitespace
characters.
\item
The constructs $q \cdot q'$, $q \sep q'$ and $q^*$ are similar to the respective
regular expression constructs. However, the regular constructs for QREs are more
restrictive for QREs since the transformations made to the data by
canonizers and lenses involve matching the data against a regular expression
deterministic. Therefore to make these transformations deterministic, the
data should be matched uniquely against the regular expressions, and this
constrains the regular expressions that may be used by QREs.
\item
The $\normalize{R}{R'}{f}$ construct allows the programmer to choose how to
canonize $\mathcal{L}(R)$ into $\mathcal{L}(R')$ be providing the canonizing
function $f$ themselves. This constuct is useful when the desired equivalence
relation cannot easily be expressed by any of the other constructs.
  \end{enumerate}

>>>>

\newcommand{\FINISH}[3]{\ifdraft\textcolor{#1}{[#2: #3]}\fi}
\newcommand{\bcp}[1]{\FINISH{dkred}{B}{#1}}
\newcommand{\BCP}[1]{\FINISH{dkred}{B}{\bf #1}}
\newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}} % Toronto Maple Leafs Blue :-)
\newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
\newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
\newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}

@@

\dpw{The next paragraph gets at the problem, but takes a while to do
so.  Why not make the first sentence go directly to the problem:  ``Unfortunately,
quotient lens programs are unnecessarily verbose and challenging to construct.'' ... or
something like that.}

@@

\saz{It might be a good idea to show what this would look like in all its gory
detail, just to make the problem very concrete. Perhaps in a figure?}\bcp{+1}\dpw{+1}

In this paper, we introduce the formalism of Quotient Regular Expressions (or
QREs) to address this issue. \dpw{cut unnecessary phrases that add little information
content.  eg, cut ``the formalism of.''  Doing so makes your text ``punchier.''}  QREs enable the programmer to express
a rich set of equivalence relations on regular expressions at the level of
syntax. QREs therefore enable the programmer to express more equivalence
relations with less effort. Also, since QREs are defined at the level of
syntax on regular expressions, QREs give the programmer more control over where
the quotienting occurs.

@@

\dpw{I think it would be a little more effective to introduce QREs with a running,
but informal, concrete example. Explain these concepts informally as you go.
Then summarize the syntax at the end.}

@@

  \begin{enumerate}
    \item 
    Every regular expression can be lifted to a QRE. When a regular expression
    $R$ is used as a QRE, then $\mathcal{L}(R)$ is assumed to be quotiented by
    the equality equivalence relation on $\mathcal{L}(R)$.
    \item
    For the QRE $R \mapsto s$, the data is described by $\mathcal{L}(R)$, but
    all members of $R$ are considered equivalent to each other, with $s \in
    \mathcal{L}(R)$ as the choice of representative. This construct is useful
    when the programmer simply wishes to ignore unnecessary characters such as
    excess whitespace e.g. 
    $$\mathtt{FIRST\_NAME} \cdot (\mathtt{WHITESPACE}^* \mapsto `` \quad ")
    \cdot \mathtt{LAST\_NAME}$$
    \item
    $\squash{R}{R'}{f}$ enables the programmer to use one of two different
    representations of the data if they can provide a way of mapping the first
    representation to the second. For instance, the first representation may be
    $$\mathtt{FIRST\_NAME} \cdot `` \quad " \cdot \mathtt{LAST\_NAME}$$
    while the second representation may be
    $$\mathtt{FIRST\_NAME} \cdot ``," \cdot \mathtt{LAST\_NAME}$$
    The mapping that maps the first representation to the second simply converts
    the space between the first and last name to a comma.
    \item
    The $\perm{q_1, \ldots, q_n}{q}$ construct enables the programmer to
    consider all permutations of some data that are interspersed with a
    separator the same. For instance, in the \textsc{Bib}\TeX{} to EndNote
    example, the programmer may use this construct as follows:
    $$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
    \re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
 where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$ are
regular expressions describing the respective \textsc{Bib}\TeX\; fields of the
same name, and $\re{WHITESPACE}$ is a regular expression describing whitespace
characters.
\item
The constructs $q \cdot q'$, $q \sep q'$ and $q^*$ are similar to the respective
regular expression constructs. However, the regular constructs for QREs are more
restrictive for QREs since the transformations made to the data by
canonizers and lenses involve matching the data against a regular expression
deterministic. Therefore to make these transformations deterministic, the
data should be matched uniquely against the regular expressions, and this
constrains the regular expressions that may be used by QREs. \dpw{This is not an
``example'' -- it is an informal description of how these operators work.}
\item
The $\normalize{R}{R'}{f}$ construct allows the programmer to choose how to
canonize $\mathcal{L}(R)$ into $\mathcal{L}(R')$ be providing the canonizing
function $f$ themselves. This constuct is useful when the desired equivalence
relation cannot easily be expressed by any of the other constructs.
  \end{enumerate}
